---
title: "DATA-613-001-Summer Data Science"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1. Anna Hyunjung Kim

## Module 2: Git and GitHub

This assignment will Consider the data in the nycflights13 package.

-   **Problem 2**: Is there a relationship between the age of a plane and its delays?

-   **Problem 3**: Find the 10 days of the year that have the highest median departure delay, then select all flights from those 10 days.

```{r}
# Install pacman if it's not already installed
if (!require("pacman")) install.packages("pacman")
```

Load the packages we need.

```{r}
# Use pacman to install (if needed) and load the packages
pacman::p_load(pacman, tidyverse,
               nycflights13)
options(dplyr.summarise.inform = FALSE)
```

```{r}
help(package = "nycflights13")
data(package = "nycflights13")

```

We can see there are five data sets.

```         
airlines                         Airline names.
airports                         Airport metadata
flights                          Flights data
planes                           Plane metadata.
weather                          Hourly weather data
```

## Is there a relationship between the age of a plane and its delays?

To explore the relationship between plane age and departure delays, I tried several approaches.

First, I made a scatterplot using all flights, but the data was too noisy to see any clear pattern. Then I calculated the correlation between plane age and delay, but it was weak and not very helpful.

So, I decided to group the data by plane (tailnum) and compute the average delay for each plane.

Lastly, I tried to group it by year and claculte it. This made the pattern easier to see.\

**The final plot shows a new pattern: older planes tended to have less delays. The trend line shows a slight negative relationship between age and delay.**

*\[I thought it's better to show the whole process because it is homework, so the results don't looks good but the final chart shows it has correlation\]*

## 2.0 Cleaning Data

```{r}
#View(planes)
## ?planes

planes %>% 
  select(tailnum, made_year = year ) -> planes_year
```

```{r}
## ?flights
## View(flights)

flights_year <- flights %>%
  left_join(planes_year, by = "tailnum")

flights_year %>% 
  mutate(plane_age = year - made_year) -> df1

head(df1)
colnames(df1)

colSums(is.na(df1))
colMeans(is.na(df1))
```

When I check the NA, dep_delay's percentage is 2.45 and arr_delay's is 2.8. So, I will remove the NA

```{r}
df_clean <-df1 %>%
  filter(!is.na(dep_delay), !is.na(arr_delay), !is.na(plane_age))

#colMeans(is.na(df_clean))
```

Now we are ready to analyse with data named 'df_clean'

## 2.1 Scatter plots

### Relationship between Plane Age and Departure Delay

First try is scatter plot. It doesn't look good and hard to find pattern or linear.

```{r}
head(df_clean)

df_clean %>% 
  ggplot(data = df_clean, mapping = aes(x = plane_age, y = dep_delay))+
  geom_point()+
    labs(title = "Relationship between Plane Age and Departure Delay",
       x = "Plane Age (years)",
       y = "Departure Delay (minutes)")

```

Next try is box plot. It seems a lot of outlier. Also, it's hard to read.

```{r}
ggplot(df_clean, mapping = aes(x = as.factor(plane_age), y = dep_delay)) +
  geom_boxplot()+
  coord_flip()+   
  labs(title = "Relationship between Plane Age and Departure Delay",
       x = "Plane Age (years)",
       y = "Departure Delay (minutes)")
```

So, then let's try to use number. I can try to use Pearson correlation coefficient using the `cor()` function. The result is -.015. It is weak correlation.

```{r}
cor(df_clean$plane_age, df_clean$dep_delay, use = "complete.obs")
```

### Relationship between Plane Age and Arrival Delay

I will try to do the same with departure delay. It also doesn't look good to find any pattern.

```{r}
df_clean %>% 
  ggplot(data = df_clean, mapping = aes(x = plane_age, y = arr_delay))+
  geom_point()+
    labs(title = "Relationship between Plane Age and Arrival Delay",
       x = "Plane Age (years)",
       y = "Arrival Delay (minutes)")

ggplot(df_clean, mapping = aes(x = as.factor(plane_age), y = arr_delay)) +
  geom_boxplot()+
  coord_flip()+   
  labs(title = "Relationship between Plane Age and Arrival Delay",
       x = "Plane Age (years)",
       y = "Arrival Delay (minutes)")
```

It's hard to say the relationship has linear. because -0.017 is not sufficient to say.

```{r}
cor(df_clean$plane_age, df_clean$arr_delay, use = "complete.obs")
```

So, I can't find the relationship between flight age and delay. But, I would like to try one more with mean of delay for each flight tailnum.

## 2.2 Second try with mean of delay per tailnum. 

Let's do departure delay first.

```{r}
df_by_tailnum <- df_clean %>%
  select(tailnum, dep_delay, plane_age) %>% 
  group_by(tailnum) %>%
  summarize(mean_dep_delay = mean(dep_delay), plane_age = first(plane_age))

df_by_tailnum 
```

```{r}
ggplot(df_by_tailnum, aes(x = plane_age, y = mean_dep_delay)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")+  
  labs(title = "Relationship between Plane Age and Departure Delay",
       x = "Plane Age (years)",
       y = "Average of Departure Delay of tailnum(minutes)")
```

```{r}
df_by_tailnum_arr <- df_clean %>%
  select(tailnum, arr_delay, plane_age) %>% 
  group_by(tailnum) %>%
  summarize(mean_arr_delay = mean(arr_delay), plane_age = first(plane_age))

df_by_tailnum_arr
```

```{r}
ggplot(df_by_tailnum_arr, aes(x = plane_age, y = mean_arr_delay)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")+  
  labs(title = "Relationship between Plane Age and Arrival Delay",
       x = "Plane Age (years)",
       y = "Average of Arrival Delay of tailnum(minutes)")
```

We can find there is no relationship between mean of delay for each flight tailnum and delay.

## 2.3 Third try with mean of delay per year.

Departure delay will be analysed first.

```{r}
df_by_tailnum_age <- df_clean %>%
  select(tailnum, dep_delay, arr_delay, plane_age) %>% 
  group_by(plane_age) %>%
  summarize(mean_dep_delay = mean(dep_delay), mean_arr_delay = mean(arr_delay), plane_age = first(plane_age))

df_by_tailnum_age
```

```{r}
ggplot(df_by_tailnum_age, aes(x = plane_age, y = mean_dep_delay)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")+  
  labs(title = "Relationship between Plane Age and Departure Delay",
       x = "Plane Age (years)",
       y = "Average of Departure Delay of tailnum(minutes)")
```

```{r}
ggplot(df_by_tailnum_age, aes(x = plane_age, y = mean_arr_delay)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")+  
  labs(title = "Relationship between Plane Age and Arrival Delay",
       x = "Plane Age (years)",
       y = "Average of Arrival Delay of tailnum(minutes)")
```

```{r}
cor(df_by_tailnum_age$plane_age, df_by_tailnum_age$mean_dep_delay, use = "complete.obs")

cor(df_by_tailnum_age$plane_age, df_by_tailnum_age$mean_arr_delay, use = "complete.obs")
```

To better understand the relationship, I use cor function the correlation between plane age and delays.

-   Correlation between plane age and mean departure delay: -0.45

-   Correlation between plane age and mean arrival delay: -0.43

It suggest that older planes tend to have fewer delays on average.

## Find the 10 days of the year that have the highest median departure delay, then select all flights from those 10 days.

### 3.0 Find the 10 days of the year that have the highest median departure delay

```{r}
top10_dep_delay <- df_clean %>%
  select(year, month, day, dep_delay) %>% 
  group_by(year, month, day) %>% 
  summarise(median_dep_delay = median(dep_delay)) %>% 
  arrange(desc(median_dep_delay)) %>% 
  head(10)

top10_dep_delay
```

```{r}
df_clean %>% 
  inner_join(top10_dep_delay, by = c("year", "month", "day")) %>% 
  arrange(desc(median_dep_delay))
  
```

## 3.1 Select all flights from those 10 days.

```{r}
top10_dep_delay %>% 
  inner_join(df_clean, by = c("year", "month", "day")) %>% 
  arrange(desc(median_dep_delay))
```
